Predicting the PCG PSeudo-Random Number Generator In Practice

### 2020-05-25 article/main.tex
algos + texte

\documentclass[preprint,svgnames]{iacrtrans}

\usepackage[utf8]{inputenc}
\usepackage{amssymb, amsmath, amsfonts, amscd}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{subcaption}
\usepackage{algorithm2e}
\usepackage{tikz}
\usepackage{cellspace}
\usepackage{multirow}
%\usepackage{parskip}
\usetikzlibrary{patterns}

@@

\usepackage[pdftex,bookmarks,bookmarksopen,bookmarksdepth=3]{hyperref}
\hypersetup{colorlinks=true,citecolor=red,linkcolor=red,urlcolor=black}

\DeclareMathSymbol{:}{\mathord}{operators}{"3A}

\usepackage{algorithmic}

% knuth-style algos
\newcommand{\slug}{\hbox{\kern1.5pt\vrule width2.5pt height6pt depth1.5pt\kern1.5pt}}
\def\xskip{\hskip 7pt plus 3pt minus 4pt}
\newdimen\algindent
\newif\ifitempar \itempartrue % normally true unless briefly set false
\def\algindentset#1{\setbox0\hbox{{\bf #1.\kern.25em}}\algindent=\wd0\relax}
\def\algbegin #1 #2{\algindentset{#21}\alg #1 #2} % when steps all have 1 digit
\def\aalgbegin #1 #2{\algindentset{#211}\alg #1 #2} % when 10 or more steps
\def\aaalgbegin #1 #2{\algindentset{#2111}\alg #1 #2} % when 10 or more steps
\def\alg#1(#2). {\medbreak % Usage: \algbegin Algorithm A (algname). This...
  \noindent{\bf#1}({\it#2\/}).\xskip\ignorespaces}
\def\algstep#1.{\ifitempar\smallskip\noindent\else\itempartrue
  \hskip-\parindent\fi
  \hbox to\algindent{\bf\hfil #1.\kern.25em}%
  \hangindent=\algindent\hangafter=1\ignorespaces}
% end of borrowed macros

@@

This section describes the \textsf{PCG64} non-cryptographic pseudo-random number
generator (more precisely \textsf{PCG-XSL-RR} in the designer's terminology). It
has an internal state of 128-bit, which operate as a linear congruential
generator modulo~$2^{128}$. More precisely:
\[
  S_{i+1} = a S_i + c \bmod 2^{128},
\]
Where $a = 47026247687942121848144207491837523525$ is a fixed 126-bit
constant. The first initial state $S_0$ is the seed of the generator. The
increment $c$ may be specified by the user of the PRNG to produce different
output streams with the same seed (just as the IV acts in a stream cipher). If
no value of $c$ is specified, then a default increment is provided:
$c = 117397592171526113268558934119004209487$.

@@

@@

\section{Warm-up: A Prediction Algorithm for \textsf{PCG64} With Known Increment}

@@

In this case, reconstructing the 128-bit internal state $S_i$ of the generator
is sufficient to produce the pseudo-random flow with 100\% accuracy (the
generator can also be clocked backwards if necessary, so that the seed can be
easily reconstructed). We therefore focus on reconstructing $S_0$ (the seed)
from $X_0, X_1, X_2, \dots$. A very simple strategy could be the following:
\begin{enumerate}
\item Guess the 64 upper bits of $S_0$.
\item Compute the missing 64 lower bits using~\eqref{eq:output}, with:
\[
   S_0[0:64] = S_0[64:128] \oplus (X_0  \lll S[122:128]).
\]
\item If $X_1$ is computed correctly, output $S_0$.
\end{enumerate}

This requires $2^{64}$ iterations of a loop that does a dozen arithmetic
operations. This is the baseline that we improve upon. An improved
``guess-and-determine'' prediction algorithm is possible, which essentially
amounts to expose a truncated version of the underlying linear congruential
generator, and attack it using the well-known tools. This is possible by
combining the following ingredients:
\begin{itemize}
\item The underlying linear congruential generator uses a power-of-two modulus,
  therefore the $\ell$ low-order bits of $S_{i+1}$ are entirely determined by
  the $\ell$ low-order bits of $S_i$. More precisely, we have:
  \begin{equation}\label{eq:lcg}
    S_{i+1} = aS_i + c \bmod 2^\ell, \qquad \text{for all } 0 \leq \ell \leq 128
  \end{equation}
  Therefore, guessing low-order bits of $S_0$ yields a ``long-term advantage''
  that holds for all subsequent states.

@@

In figure~\ref{fig:Cknown}, we see that guessing $S_0[0:\ell]$ and a few 6-bit
rotations give access to $S_i[58:64+\ell]$ for the corresponding
states. Therefore, looking at $S_i[\ell:64+\ell]$, we are facing a truncated
linear congruential generator on $64$ bits, where we have access to the most
$6+\ell$ bits of each state, for a few consecutive states. This is sufficient to
reconstruct entirely the successive states of this truncated linear congruential
generator. This reveals $S_0[\ell:64+\ell]$, and using~\eqref{eq:output} the
entire $S_0$ can be reconstructed. The precise details follow.

@@

  % T_0
  \begin{scope}[xshift=4.6cm, yshift=-1.5cm]    
    \draw[dashed]  (0, 0) rectangle +(6.4, 1);
    \draw[thick]  (0, 0) rectangle +(2.4, 1);
    \draw[]  (1.8, 0) rectangle +(0, 1);
    \path (2.4, 0) rectangle node {$???????? \cdots ????????$} (6.4, 1);

@@

  % T_1
  \begin{scope}[xshift=4.6cm, yshift=-4.5cm]    
    \draw[dashed]  (0, 0) rectangle +(6.4, 1);
    \draw[thick]  (0, 0) rectangle +(2.4, 1);
    \draw[]  (1.8, 0) rectangle +(0, 1);
    \path (2.4, 0) rectangle node {$???????? \cdots ????????$} (6.4, 1);

@@

  % T_2
  \begin{scope}[xshift=4.6cm, yshift=-7.5cm]    
    \draw[dashed]  (0, 0) rectangle +(6.4, 1);
    \draw[thick]  (0, 0) rectangle +(2.4, 1);
    \draw[]  (1.8, 0) rectangle +(0, 1);
    \path (2.4, 0) rectangle node {$???????? \cdots ????????$} (6.4, 1);

@@

Set $S_i' = S_i - S_0[0:\ell] a^i - c \sum_{j=0}^{i-1} a^i$. The sequence
$(S'_i)$ describe the internal states we would observe if $c=0$ and
$S_0[0:\ell] = 0$. It is easy to check that $S'_{i} = a^i (S_0 - S_0[0:\ell])$,
and therefore the successive $S'_i$'s follow a geometric progression. % Indeed :
% \begin{align*}
% S_i &= a S_{i-1} + c\\
%     &= a^2 S_{i-2} + (a + 1) c \\
%     &\vdots  \\
%     &= a^i S_0  + c \sum_{j=0}^{i-1} a^i \\
%   S'_i  &= S_i - c \sum_{j=0}^{i-1} a^i - S_0[0:\ell] a^i \\
%     &= a^i (S_0 - S_0[0:\ell])
% \end{align*}
The $\ell$ low-order bits of $S_0 - S_0[0:\ell]$ are all zero by construction;
therefore set $T_i = \left(S'_i / 2^\ell\right) \bmod 2^{64}$ (the division by
$2^\ell$ is exact). It follows that the $(T_i)$ is also a geometric progression
with common ratio $a$, and we have access to the $\ell+6$ high order bits of
$T_i$ if we guess the $i$-th rotation. We have seen in
section~\ref{sec:geometric} that 19 known bits on 3 consecutive outputs are
sufficient for a deterministic reconstruction of the full sequence. Therefore,
we guess 19 low-order bits of the state and three consecutive rotations.

@@

\begin{center}\noindent\framebox{
  \begin{minipage}{0.95\textwidth}
    \algbegin Algorithm K (State reconstruction with known
    increment). Reconstruct $S_0$ from three consecutive outputs $X_j$
    ($0 \leq j < 3$). In all steps of the algorithm, we have $j= 0, 1, 2$.

\algstep K1. [Guess state] For each $0 \leq w < 2^{19}$, set $w_j = a^j w \bmod 2^{19}$ and do steps K2--K0.

\algstep K2. [Guess rotations] For each $0 \leq r_0, r_1, r_2 < 64$, do steps K3--K9.

\algstep K3. [Undo rotations] Set $Y_j \gets X_j \lll r_j$. % ($Y_j$ is the $j$-th output on which the rotation has been undone).

\algstep K4. [Build $T$] Set $T_j \gets \left(r_j \oplus Y_j[58:64]\right) \oplus \bigl( 64 \cdot \left(w_j \oplus Y_j[0:19]\right)\bigr)$.

\algstep K5. [Shift $T$] Set $T'_j \gets  T_j - (a^j w + c \sum_{k = 0}^j a^k)[58:83]$

\algstep K6. [Truncated LCG] Set $(U_0, U_1, U_2) \gets \left\lfloor (T_0, T_1, T_2) \cdot 2^{39} \cdot \widetilde G^{-1} \right\rceil \cdot \widetilde G$.

\algstep K7. [Reconstruct $S_0$] Set $\widehat S_0[0:64] \gets w + 2^{19} \cdot U_0[0:45]$ and $\widehat S_0[64:128] \gets S_0[0:64] \oplus Y_0$.

\algstep K8. [Recompute $Y_1$] Set $\widehat S_1 \gets a \widehat S_0 + c$ and $\widehat Y_1 = \widehat S_1[0:64] \oplus \widehat S_1[64:128]$.

\algstep K9. [Check] If $\widehat{Y}_1 = Y_1$, then output $\widehat S_0$ as a candidate internal state.
\end{minipage}
}
\end{center}

\subsection{preuve cas facile}

@@

	Hence, the smallest l such that the inequality is satisfied is \(\ell = 8\) %7 avec borne CVP
\end{itemize}

[insert here] attack predicting algo description.

\section{Predicting PCG with Unknown Increment}
We can't use the same algorithm for the general case. We don't known $C$, so we used an other geometric reduction.
$(\Delta S_{i+1}S_i)_i$ est un suite géométrique de raison $A$.

On calcule préalablement les matrices $G$ et $G^{-1}$.

@@

\subsection{Part 1 : Premier filtre}
Pour tous les $r_i$, les $S_0[0 : \ell]$ et les $C[0 : \ell]$ possibles:
\begin{enumerate}
  \item On récupère le $T$  correspondant à nos suppositions. On calcule $\Delta T$ :
  \begin{equation}
    \Delta T_i = (S_{i+1}[58 : 64 + \ell] - S_i[58 : 64 + \ell] - (S_0[0:\ell] * (A^{i+1} - A^i) + C[0:\ell] * A^i )[58 : 64 + \ell]) * 2^{64 - l - 6}
  \end{equation}
  Ou plus simplement $\Delta T_i = T'_{i+1} - T'_{i}$ avec $T'$ dont on aurait soustrait la composante en $C[0:\ell]$ et non en $C$.

>>>>

\documentclass[nohyperref,preprint,svgnames]{iacrtrans}

\usepackage[utf8]{inputenc}
\usepackage{amssymb, amsmath, amsfonts, amscd}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{subcaption}
%\usepackage{algorithm2e}
\usepackage{tikz}
\usepackage{cellspace}
\usepackage{multirow}
%\usepackage{parskip}
\usetikzlibrary{patterns}

@@

%\usepackage[pdftex,bookmarks,bookmarksopen,bookmarksdepth=3]{hyperref}
%\hypersetup{colorlinks=true,citecolor=red,linkcolor=red,urlcolor=black}

\DeclareMathSymbol{:}{\mathord}{operators}{"3A}
\DeclareMathOperator{\LCG}{LCG}
\DeclareMathOperator{\MCG}{MCG}

\usepackage[noend]{algpseudocode}
%\usepackage{algorithmicx}

% knuth-style algos
% \newcommand{\slug}{\hbox{\kern1.5pt\vrule width2.5pt height6pt depth1.5pt\kern1.5pt}}
% \def\xskip{\hskip 7pt plus 3pt minus 4pt}
% \newdimen\algindent
% \newif\ifitempar \itempartrue % normally true unless briefly set false
% \def\algindentset#1{\setbox0\hbox{{\bf #1.\kern.25em}}\algindent=\wd0\relax}
% \def\algbegin #1 #2{\algindentset{#21}\alg #1 #2} % when steps all have 1 digit
% \def\aalgbegin #1 #2{\algindentset{#211}\alg #1 #2} % when 10 or more steps
% \def\aaalgbegin #1 #2{\algindentset{#2111}\alg #1 #2} % when 10 or more steps
% \def\alg#1(#2). {\medbreak % Usage: \algbegin Algorithm A (algname). This...
%   \noindent{\bf#1}({\it#2\/}).\xskip\ignorespaces}
% \def\algstep#1.{\ifitempar\smallskip\noindent\else\itempartrue
%   \hskip-\parindent\fi
%   \hbox to\algindent{\bf\hfil #1.\kern.25em}%
%   \hangindent=\algindent\hangafter=1\ignorespaces}
% end of borrowed macros

@@

This section describes the \textsf{PCG64} non-cryptographic pseudo-random number
generator (more precisely \textsf{PCG-XSL-RR} in the designer's terminology). It
has an internal state of 128-bit, which operate as a linear congruential
generator modulo~$2^{128}$. More precisely:
\[
  S_{i+1} = a S_i + c \bmod 2^{128},
\]
Where $a = 47026247687942121848144207491837523525$ is a fixed 126-bit
constant. The first initial state $S_0$ is the seed of the generator. The
increment $c$ may be specified by the user of the PRNG to produce different
output streams with the same seed (just as the IV acts in a stream cipher). If
no value of $c$ is specified, then a default increment is provided:
$c = 117397592171526113268558934119004209487$. Note that $c$ must be odd.

@@

\subsection{Reconstructing Linear Congruential Generators}

Given a bit-size $k$, a fixed multiplier $a$ and an increment $c$, define the
sequence $u_0 = x, u_{i+1} = a u_i + c \bmod 2^k$. The sequence $u$ form the
successive internal states of a linear congruential generator (LCG). When the
increment is zero, the sequence forms a geometric progression with common ratio
$a$, and is sometimes referred to as the output of a multiplicative congruential
generator (MCG).

Let $\LCG_{n,k}(x, c)$ denote the vector
$(u_0, \dots, u_{n-1}) \in \mathbb{Z}_{2^k}$ (we will omit $n$ and $k$ when they
are obvious from the context).  It is easy to check that:
\begin{align}
  \label{eq:lcg-additive}
  \LCG_{n,k}(x + y, c + d) &= \LCG_{n,k}(x, c) + \LCG_{n,k}(y, d),  \\
  \label{eq:lcg-scalar}
  \LCG_{n,k}(\lamba x, \lambda c) = \lambda \LCG_{n,k}(x, c).

\end{align}

@@

\section{Warm-up: A Prediction Algorithm for \textsf{PCG64} With Known Increment}
\label{sec:Cknown}

@@

In this case, reconstructing the 128-bit internal state $S_i$ of the generator
is sufficient to produce the pseudo-random flow with 100\% accuracy (the
generator can also be clocked backwards if necessary, so that the seed can be
easily reconstructed). We therefore focus on reconstructing $S_0$ (the seed)
from $X_0, X_1, X_2, \dots$. A very simple strategy could be the following:
\begin{enumerate}
\item Guess the 64 upper bits of $S_0$.
\item Compute the missing 64 lower bits using~\eqref{eq:output}, with:
\[
   S_0[0:64] = S_0[64:128] \oplus (X_0  \lll S[122:128]).
\]
\item If $X_1$ is computed correctly, then output $S_0$.
\end{enumerate}

This ``baseline'' procedure requires $2^{64}$ iterations of a loop that does a
dozen arithmetic operations; it always output the correct value of $S_0$, and
may output a few other ones (they can be easily discarded by checking $X_2$). An
improved ``guess-and-determine'' prediction algorithm is possible, which
essentially amounts to expose a truncated version of the underlying linear
congruential generator, and attack it using the well-known tools. This is
possible by combining the following ingredients:
\begin{itemize}
\item The underlying linear congruential generator uses a power-of-two modulus,
  therefore the $\ell$ low-order bits of $S_{i+1}$ are entirely determined by
  the $\ell$ low-order bits of $S_i$. More precisely, we have:
  \begin{equation}\label{eq:lcg}
    S_{i+1} = aS_i + c \bmod 2^\ell, \qquad \text{for all } 0 \leq \ell \leq 128
  \end{equation}
  Therefore, guessing low-order bits of $S_0$ yields a ``long-term advantage''
  that holds for all subsequent states.

@@

In figure~\ref{fig:Cknown}, we see that guessing $S_0[0:\ell]$ and a few 6-bit
rotations give access to $S_i[58:64+\ell]$ for the corresponding
states. Therefore, looking at $S_i[\ell:64+\ell]$, we are facing a truncated
linear congruential generator on $64$ bits, where we have access to the most
$6+\ell$ bits of each state (denoted by $T$), for a few consecutive states. This
is sufficient to reconstruct entirely the successive states of this truncated
linear congruential generator. This reveals $S_0[\ell:64+\ell]$, and
using~\eqref{eq:output} the entire $S_0$ can be reconstructed. The precise
details follow.

@@

  % T_0
  \begin{scope}[xshift=4.6cm, yshift=-1.5cm]    
    \draw[dashed]  (0, 0) rectangle +(6.4, 1);
    \draw[thick]  (0, 0) rectangle +(2.4, 1);
    \draw[]  (1.8, 0) rectangle +(0, 1);
    \path (2.4, 0) rectangle node {$???????????????????$} (6.4, 1);

@@

  % T_1
  \begin{scope}[xshift=4.6cm, yshift=-4.5cm]    
    \draw[dashed]  (0, 0) rectangle +(6.4, 1);
    \draw[thick]  (0, 0) rectangle +(2.4, 1);
    \draw[]  (1.8, 0) rectangle +(0, 1);
    \path (2.4, 0) rectangle node {$???????????????????$} (6.4, 1);

@@

  % T_2
  \begin{scope}[xshift=4.6cm, yshift=-7.5cm]    
    \draw[dashed]  (0, 0) rectangle +(6.4, 1);
    \draw[thick]  (0, 0) rectangle +(2.4, 1);
    \draw[]  (1.8, 0) rectangle +(0, 1);
    \path (2.4, 0) rectangle node {$???????????????????$} (6.4, 1);

@@

We consider the first three internal states
$\mathbf{S} = (S_0, S_1, S_2) = \LCG_{3, 128}(S_0, c)$. We will guess the $\ell$
least-significant bits of $S_0$, therefore let us assume that their value is
known and denote it by $w$. We define $\mathbf{S}' = \LCG_{3, 128}(S_0 - w,
0)$. By~\eqref{eq:lcg-additive}, we have
$\mathbf{S}' = \mathbf{S} - \LCG_{3, 128}(w, c)$. The point is that the elements
of $\mathbf{S}'$ follow a geometric progression of common ratio $a$; in
addition, the $\ell$ least significant bits of each components are zero. The
crux of the reconstruction algorithm is to find $\mathbf{S}'$.

For this, set $\mathbf{T}' = \left(\mathbf{S}' / 2^\ell\right) \bmod 2^{64}$
(the division by $2^\ell$ is exact). It follows that the $\mathbf{T}'$ is also a
geometric progression with common ratio $a$, and we have access to the $\ell+6$
high order bits of $T'_i$ if we guess the $i$-th rotation. We have seen in
section~\ref{sec:geometric} that 19 known bits on 3 consecutive outputs are
sufficient for a deterministic reconstruction of the full sequence. Therefore,
we guess $\ell = 19$ low-order bits of the state and three consecutive
rotations.

@@

  \begin{algorithmic}[1]
    \Procedure{ReconstructState${}_\ell$}{$X_0, X_1, X_2$}
    \For{$0 \leq w < 2^{\ell}$} \Comment{Guess least-significant bits of $S_0$}
    \State $w_j = a^j w \bmod 2^{\ell}$ \Comment{Propagate to $S_1, S_2$}
    \State $S'_j \gets a^j w + c \sum_{k = 0}^j a^k \bmod 2^{128}$    \Comment{Shifted states}

    \For{$0 \leq r_0, r_1, r_2 < 64$} \Comment{Guess rotations}

    \State $Y_j \gets X_j \lll r_j$ \Comment{Undo rotations}

    \State $ T_j \gets \left(r_j \oplus Y_j[58:64]\right) +  64 \cdot \left(w_j \oplus Y_j[0:\ell]\right)$  \Comment{Truncated LCG output}

    \State $ T'_j \gets    T_j -  S'_i[58:64+\ell]$  \Comment{Shifted Truncated LCG output}

    \State $(U_0, U_1, U_2) \gets \left\lfloor (T'_0, T'_1, T'_2) \cdot 2^{58-\ell} \cdot \widetilde G_3^{-1} \right\rceil \cdot \widetilde G_3$ \Comment{CVP via Babai rounding}

    \State $ S_0[0:64] \gets  S'_0 + 2^{\ell} \cdot U_0[0:64-\ell]$ \Comment{Reconstruct $S_0$}
    \State $ S_0[64:128] \gets S_0[0:64] \oplus Y_0$

    \State $ S_1 \gets a  S_0 + c$ \Comment{Check consistency of $X_1$}
    \State $\widehat{Y}_1 =  S_1[0:64] \oplus  S_1[64:128]$

    \If{$\widehat {Y}_1 = Y_1$}
    \State output $ S_0$ as a candidate internal state.
    \EndIf
    \EndFor
    \EndFor
    \EndProcedure
  \end{algorithmic}

\begin{theorem}
  Algorithm \textsc{ReconstructState} is correct (it always return the correct internal state).
\end{theorem}

\begin{proof} \todo{FIX ME!}
The vector \(S'\) has been constructed such that it is in \(L(A,2^{128},n)\) and its \(\ell\) lower bits are zero.

@@

	Hence, the smallest l such that the inequality is satisfied is \(\ell = 8\) %7 avec borne CVP
\end{itemize}
\end{proof}

\section{Predicting PCG with Unknown Increment}
\label{sec:Cunknown}

The same algorithm does not apply directly to the general case where the value
of $c$ is unknown. The global strategy to obtain a prediction algorithm is
essentially the same (target a truncated version of the underlying linear
congruential generator), but is a bit more complicated to implement.

Set $\Delta S_i = S{i+1} - S_i$; it is easily checked that $\Delta S_i$ is a
geometric progression of common ratio $a$. Therefore, reconstructing both $S_0$
and $\Delta S_0$ is sufficient to compute all subsequent states (and recover the
unknown increment $c$). The prediction algorithm proceeds in three phases:
\begin{enumerate}
\item Reconstruct $\Delta S_0[0:64+\ell]$ from $X_0, \dots, X_4$.
\item Fully reconstruct $\Delta S_0$ from this partial knowledge and $X_0, \dots, X_{39}$.
\item Reconstruct $S_0$ from $\Delta S_0$ and $X_0, \dots, X_{32}$.
\end{enumerate}

\noindent Only the first phase is computationally intensive.

\subsection{Phase 1: Partial Difference Reconstruction}

In order to access to a part of $\Delta S_i$, we use the same strategy as in
section~\ref{sec:Cknown}: we guess the least significant bits of $S_0$ and some
rotations. The difference is that, in addition, to obtain the ``long-term
advantage'' in the least significant bits of the states, we must also guess the
least significant bits of $c$ as well. We must also guess one additional
rotation to compute the differences $\Delta S_i$.

Another problem is that of confirming that the guesses are correct. When $c$ is
known, we could reconstruct the internal state; from there, filtering out the
bad guesses is easy. When $c$ is unknown, a very strong consistency check can be
implemented as follows: once we have found $\Delta S_0[0:64+\ell]$, we can
compute $\Delta S_i[0:64+\ell]$ for any $i$; because we have guessed the first
rotation and $\ell$ lest significant bits of the state, using~\eqref{eq:output}
we gain access to $S_0[58:64+\ell]$; combined with the ``differences'', this
reveals $S_i[58:64+\ell]$ for any $i$ (and we already had $S_i[0:\ell]$). This
means that we can compute $S_i[0:\ell] \oplus S_i[64:64+\ell]$ for any $i$.

Given a ``fresh'' output $X_i$, we can enumerate the 64 possible rotations
$r_i$; for each one, we can check whether
$(X_i \lll r_i)[0:\ell] = S_i[0:\ell] \oplus S_i[64:64+\ell]$. If none of the
rotations yields a match, then the guesses made beforehand have to be wrong. As
soon as $\ell \geq 6$, this test is efficient (and it can be done for any number
of fresh $X_i$'s). As a consequence, bad guesses can be filtered deterministically.

\begin{align*}
  \mathbf{S} &= \LCG(S_0, c) \\
             &= \LCG\left(S_0[0:\ell] + 2^\ell S_0[\ell:128], c[0:\ell] + 2^\ell c[\ell:128]\right) \\
             &= \LCG\left(S_0[0:\ell], c[0:\ell]) + 2^\ell \LCG\left(S_0[\ell:128],  c[\ell:128]\right) \\
  \Delta \mathbf{S} &= \LCG(S_1, c) - \LCG(S_0, c) \\
             &= \LCG(S_1 - S_0, 0)
\end{align*}

\begin{algorithmic}[1]
  \Procedure{ReconstructPartialDifference}{$X_0, \dots, X_9$}
  \For{$0 \leq w_0 < 2^{\ell}$ and $0 \leq c < 2^{\ell - 1}$} \Comment{Guess least-significant bits}
  \State $S'_j \gets a^j w_0 + (2c+1) \sum_{k = 0}^j a^k \bmod 2^{128}$ \Comment{Shifted states}
  \For{$0 \leq r_0, r_1, r_2, r_3, r_4 < 64$} \Comment{Guess rotations}
  \State $Y_j \gets X_j \lll r_j$ \Comment{Undo rotations}
  \State $T_j \gets \left(r_j \oplus Y_j[58:64]\right) +  64 \cdot \left(w_j \oplus Y_j[0:\ell]\right)$ \Comment{Truncated LCG}
  \State $T'_j \gets T_j -  S'_i[58:64+\ell]$ \Comment{Shifted truncated LCG}
  \State $\Delta T'_j = T'_{j+1} - T'_j$  \Comment{Difference}
  \State $(\Delta U_0, \Delta U_1, \Delta U_2, \Delta U_3) \gets \left\lfloor (\Delta T'_0, \Delta T'_1, \Delta T'_2, \Delta T'_3) \cdot 2^{39} \cdot \widetilde G_4^{-1} \right\rceil \cdot \widetilde G_4$ \Comment{CVP}
  \State $\Delta S_0[0:64+\ell] \gets \left((w_1 - w_0) \bmod 2^{\ell}\right) + 2^{\ell} \cdot U_0[0:58-\ell]$
  \State ...
  \EndFor
  \EndFor
  \EndProcedure
\end{algorithmic}

@@

\subsection{Part 1 : Premier filtre}

Pour tous les $r_i$, les $S_0[0 : \ell]$ et les $C[0 : \ell]$ possibles:
\begin{enumerate}
  \item On récupère le $T$  correspondant à nos suppositions. On calcule $\Delta T$ :
  \begin{equation}
    \Delta T_i = (S_{i+1}[58 : 64 + \ell] - S_i[58 : 64 + \ell] - (S_0[0:\ell] * (A^{i+1} - A^i) + C[0:\ell] * A^i )[58 : 64 + \ell]) * 2^{64 - l - 6}
  \end{equation}
  Ou plus simplement $\Delta T_i = T'_{i+1} - T'_{i}$ avec $T'$ dont on aurait soustrait la composante en $C[0:\ell]$ et non en $C$.
